\documentclass{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{xspace}
\usepackage{needspace, etoolbox}
\usepackage{url}

\usepackage{mdframed}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{thm-restate}

\usepackage{tikz}
\usepackage{tikz-cd}
\usetikzlibrary{fit, backgrounds, calc}


\input{macros}

\title{Finitely Descriptible Properties}
\author{
Clément Michaud \\
Independent Researcher\\
\texttt{clement.michaud34@gmail.com}
}
\date{March 2025}

\begin{document}

\maketitle

\begin{center}
\begin{minipage}{0.9\textwidth}
\small
\textbf{Abstract.} In mathematics, objects are often exhibited with special properties that can be characterized by universal properties in the language of category theory. This paper analyzes the finitary nature of the transformations between such objects and proposes a framework for encoding mathematics purely through graphs, maintaining their categorical structure and universal properties.
\end{minipage}
\end{center}

\vspace{10pt}


\section{Introduction}

Mathematical objects typically arise as particular instances of universal constructions, where transformations between them preserve structure in a finitary manner. In this paper, we study how these transformations, being finitely describable, can be represented within a purely graphical framework, enabling formal reasoning in computational settings.

\section{Preliminaries}
\subsection{Categories and Universal Properties}

Category theory provides a general framework to study mathematical structures and the relationships between them. At its core lies the notion of a category, which abstracts the idea of objects and the arrows that connect them.


\begin{definition}[Category]
A \emph{category} is a collection of \emph{objects} linked by \emph{arrows} (also called \emph{morphisms}). A category has two basic properties:
\begin{itemize}
    \item The ability to compose arrows associatively.
    \item The existence of an identity arrow for each object.
\end{itemize}
We denote the set of objects of a category $\mathcal{C}$ by $\operatorname{Ob}(\mathcal{C})$ and the set of morphisms by $\operatorname{Mor}(\mathcal{C})$.
\end{definition}

Beyond merely listing objects and morphisms, categories allow us to isolate particular objects through the notion of \emph{universal properties}. Universal properties are not incidental: they are fundamental to the way mathematics is developed. Every mathematical concept that we exhibit — from basic structures like groups or sets to more complex constructions — is characterized by a universal property that selects it uniquely, up to isomorphism, among all objects sharing a given structure.

For instance, when we speak of the group of integers $(\mathbb{Z}, +)$, we are selecting, among all groups, the free group generated by a single element, characterized by its universal mapping property. Similarly, when we refer to the natural numbers, we mean the unique (up to isomorphism) initial object in the category of inductive sets: a set equipped with a distinguished element (zero) and a successor function satisfying Peano's axioms. In each case, variable declarations and structural constraints are used to exhibit an object with the desired universal property, ensuring that our mathematical constructions are canonical and unambiguous.

\begin{definition}
A \emph{universal property} characterizes an object (up to unique isomorphism) by the existence and uniqueness of morphisms satisfying certain conditions relative to other objects.
\end{definition}

In a more formal categorical setting (see, for instance, \cite{MacLane}, or the exposition in \href{https://en.wikipedia.org/wiki/Universal_property}{Wikipedia}), universal properties are often expressed via diagrams and functors:

\begin{definition}[Universal Property --- Formal]
Let \(\mathcal{C}\) be a category, and let \(F \colon \mathcal{J} \to \mathcal{C}\) be a functor from a small category \(\mathcal{J}\) into \(\mathcal{C}\). An object \(U\) in \(\mathcal{C}\) equipped with a natural transformation 
\[
\alpha \colon \Delta_U \Longrightarrow F
\]
(where \(\Delta_U\) is the functor that sends every object of \(\mathcal{J}\) to \(U\) and every morphism to the identity on \(U\)) is said to satisfy a \emph{universal property} if, for every object \(X\) in \(\mathcal{C}\) with a natural transformation \(\beta \colon \Delta_X \Longrightarrow F\), there exists a unique morphism \(\phi \colon X \to U\) in \(\mathcal{C}\) such that \(\beta\) factors through \(\alpha\) via \(\phi\). Symbolically, we require that
\[
\beta = \alpha \circ (\Delta_{\phi}),
\]
and that \(\phi\) is unique with this property.

Intuitively, \(U\) is universal if any other \(\mathcal{C}\)-object \(X\) that tries to ``map into'' the diagram \(F\) must do so by factoring uniquely through \(U\).
\end{definition}

Universal properties in this sense encapsulate constructions ranging from products and coproducts (binary or otherwise) to free objects, limits, adjoints, and more. They ensure that once such an object \((U, \alpha)\) exists, it is determined up to isomorphism. By working with universal properties, we thus obtain concise, canonical descriptions of fundamental mathematical structures.

\input{ideas}

\end{document}
